---
name: "Mehmet"
surname: "Kutluay"
position: "Data Science Manager"
address: "Rotterdam - The Netherlands"
phone: +31 685478120
email: "mehkutluay@gmail.com"
github: "YMehmetKutluay"
linkedin: "mehmet-kutluay-phd"
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
```

# Skills

## Data Science

* Network Analytics, Anomaly Detection, Spark, Time Series Forecasting, Cross-Sectional Prediction,  Data Visualization, ETL, Remote Sensing

## Syntax and Programming

* Statistics/Data Science: Python, R, SQL, Matlab
* Engineering: Bash, Docker, Terraform
* Other: Markdown, Microsoft Office

## Languages

* Turkish: Native
* English: Fluent, Business Proficient
* Dutch: Fluent, Business Proficient

# Experience

```{r}
# Define information on Zanders
Zanders <- c("Data Science Manager", "Zanders", "April 2026 - Present", "Utrecht")
# Define information on Sam Media
Sam <- c("Head of Data", "Sam Media", "September 2025 - April 2026", "Amsterdam")
# Define information on Knab
Knab <- c("Lead Data Scientist", "Knab", "August 2023 - July 2025", "Amsterdam")
# Define information on Apollo Agriculture
Apollo <- c("Senior Data Scientist", "Apollo Agriculture", "May 2022 - April 2023", "Amsterdam")
# Define information on TMNL
TMNL <- c("Senior Data Scientist", "Transactie Monitoring Nederland", "May 2021 - April 2022", "Purmerend")
# Define information on FEC
ING_fec <- c("Data Scientist", "ING Bank N.V. - Financial Crime Analytics", "Nov 2019 - April 2021", "Amsterdam")
# Define information on Finance Analytics
ING_finance <- c("Data Scientist", "ING Bank N.V. - Finance Analytics", "Feb 2018 - Oct 2019", "Amsterdam")
# Define information on PhD
phd <- c("PhD Researcher in Environmental Economics", "VU University Amsterdam", "Sep 2014 - Aug 2018", "Amsterdam")
# Define information on teaching assistant
teaching_assistant <- c("Teaching Assistant", "VU University Amsterdam / Tinbergen Institute", "Sep 2013 - Feb 2018", "Amsterdam")
# Combine all information
tribble(
  ~ role, ~ company, ~ dates, ~ loc, ~ details,
  # Zanders
  Zanders[1], Zanders[2], Zanders[3], Zanders[4],
  "Building up Zanders' capabilities in appyling machine learning and Gen AI in financial institutions.",
  # Sam Media
  Sam[1], Sam[2], Sam[3], Sam[4],
  "Lead the central Data Team of 4-5 FTE's consisting of data analysts and data engineers.",
  Sam[1], Sam[2], Sam[3], Sam[4],
  "Envisioned, developed and executed product-based data strategy.",
  Sam[1], Sam[2], Sam[3], Sam[4],
  "Re-established Data Team's position within the company by re-building stakeholder relationships and deploying smarter reporting tools.",
  Sam[1], Sam[2], Sam[3], Sam[4],
  "Provided soft skills mentorship and guidance to team members.",
  # Knab
  Knab[1], Knab[2], Knab[3], Knab[4],
  "Initially joined as senior data scientist, worked as lead data scientist from March 2024.",
  Knab[1], Knab[2], Knab[3], Knab[4],
  "Lead and mentored a team of 6-7 data scientists, while also working on model development and deployment.",
  Knab[1], Knab[2], Knab[3], Knab[4],
  "Deployed first ML models for the bankd, initially in the transaction monitoring domain.",
  Knab[1], Knab[2], Knab[3], Knab[4],
  "Set up and developed the governance and technical framework for deploying more ML models. Established working relationships of data science with relevant teams (data engineering, data governance, business stakeholders and model risk management).",
    Knab[1], Knab[2], Knab[3], Knab[4],
  "Designed and executed the ML / AI strategy of Knab.",
  # Apollo
  Apollo[1], Apollo[2], Apollo[3], Apollo[4],
  "Worked in a team of 4-5 data scientists, across multiple projects with different stakeholders. Projects entailed fraud detection, credit risk scoring and satellite image database creation.",
  Apollo[1], Apollo[2], Apollo[3], Apollo[4],
  "Launched and monitored multiple models in production. Deployments involved both batch processes and real-time applications.",
  Apollo[1], Apollo[2], Apollo[3], Apollo[4],
  "Set up the data science way-of-working (Docker for local development, coding standards, created/maintained Python microservices and more).",
  Apollo[1], Apollo[2], Apollo[3], Apollo[4],
  "Optimized and strengthened existing models by incorporating unit / integration tests and documentation for ease of onboarding.",
  # TMNL
  TMNL[1], TMNL[2], TMNL[3], TMNL[4],
  "Worked in a multifunctional department that consisted of data scientists and anti-money laundering experts.",
  TMNL[1], TMNL[2], TMNL[3], TMNL[4],
  "Created/maintained two Python packages for network analytics projects: 1. efficient network traversals and 2. detecting communities within networks. Both generated sample alerts later sent to shareholder banks.",
  TMNL[1], TMNL[2], TMNL[3], TMNL[4],
  "Main contact with engineering teams to ensure a workable framework for model development. Actively helped with setting up a feature store and enforcing Spark best practices.",
  TMNL[1], TMNL[2], TMNL[3], TMNL[4],
  "Integrated anti-money laundering experts actively into advanced analytics projects.",
  TMNL[1], TMNL[2], TMNL[3], TMNL[4],
  "Participated in high-level discussions on day-to-day and strategic topics for TMNL's anti-money laundering models.",
  # ING FEC
  ING_fec[1], ING_fec[2], ING_fec[3], ING_fec[4],
  "Worked in the global team of data scientists that developed models to detect money laundering and other financial economic crimes.",
  ING_fec[1], ING_fec[2], ING_fec[3], ING_fec[4],
  "Lead technical validation of machine learning transaction monitoring models in ING Turkey.",
  ING_fec[1], ING_fec[2], ING_fec[3], ING_fec[4],
  "Created/maintained two Python packages for: 1. streamlined unsupervised anomaly detection and 2. extracting network features from transaction data.",
  ING_fec[1], ING_fec[2], ING_fec[3], ING_fec[4],
  "Machine learning domain focus: unsupervised anomal detection (e.g. isolation forest, k-means clustering, outlier variational autoencoder).",
  ING_fec[1], ING_fec[2], ING_fec[3], ING_fec[4],
  "Organized and ran workshops on day-to-day usage of Docker.",
  ING_fec[1], ING_fec[2], ING_fec[3], ING_fec[4],
  "Coached trainees to become data scientists.",
  # ING finance
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Worked in a team of 2-3 data scientists that supported employees in Group Finance. All projects were centered around time series data cleaning, wrangling and forecasting.",
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Created three R packages for time series forecasting, detecting/correcting anomalies and data exploration. These R packages are open sourced and maintained by me. Click for
  \\href{https://github.com/ing-bank/tstools}{tstools}, \\href{https://github.com/ing-bank/tsclean}{tsclean}, \\href{https://github.com/ing-bank/tsforecast}{tsforecast}",
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Machine learning domain focus: time series forecasting (ARIMA, LSTM, Kalman filter).",
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Deployed projects via remote and Docker solutions.",
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Organized and ran workshops in R, Python and statistics.",
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Coached trainees to become data scientists.",
  ING_finance[1], ING_finance[2], ING_finance[3], ING_finance[4],
    "Engaged stakeholders to help them become more data-savvy and understand what they can do effectively with data science.",
  # PhD
  phd[1], phd[2], phd[3], phd[4],
    "Introduced R as a syntax to replace licensed software (Stata) for doing statistical analyses within department.",
  phd[1], phd[2], phd[3], phd[4],
    "Organized and executed a field survey in Mumbai, India between March - June 2016. Trained a team of five surveyors and surveyed above 1500 residents.",
  # Teaching assistant
  teaching_assistant[1], teaching_assistant[2], teaching_assistant[3], teaching_assistant[4],
    "Courses: Statistics/Econometrics (masters level - including coding in SPSS and R), Mathematics (masters level - linear algebra and advanced calculus)"
  ) %>% 
  detailed_entries(role, loc, company, dates, details, .protect = F)
```

# Education

```{r}
# Define information on PhD
PhD_info <- c("PhD in Economics", "VU University Amsterdam", "Amsterdam, The Netherlands", "Sep 2014 - May 2019")
# Define information on Master
MSc_info <- c("MPhil in Economics", "Tinbergen Institute", "Amsterdam/Rotterdam, The Netherlands","Sep 2012 - Aug 2014")
# Define information on Bachelor
BSc_info <- c("BSc in Economics", "Bilkent University", "Ankara, Turkey", "Sep 2008 - Jun 2012")
# Combine all information
tribble(
    ~ degree, ~ uni, ~ loc, ~ dates, ~ details,
    # PhD
    PhD_info[1], PhD_info[2], PhD_info[3], PhD_info[4],
    "Specialization: Applied micro-econometrics, environmental economics",
    PhD_info[1], PhD_info[2], PhD_info[3], PhD_info[4],
    "Thesis title: 'The Value of (Avoiding) Malaria' (\\href{https://www.dropbox.com/s/0du7immzy1k14o1/PhD_Thesis.pdf?dl=0}{link})",
    PhD_info[1], PhD_info[2], PhD_info[3], PhD_info[4],
    "Supervisors: Roy Brouwer \\& Richard Tol",
    PhD_info[1], PhD_info[2], PhD_info[3], PhD_info[4],
    "Links to completed papers:
    \\begin{itemize}
    \\item Valuing malaria morbidity     (\\href{https://www.tandfonline.com/doi/full/10.1080/21606544.2019.1581094}{link})
    \\item Preference updating in public health risk valuation (\\href{https://www.researchgate.net/publication/320234394_Preference_updating_in_public_health_risk_valuation}{link})
    \\item Public preferences and valuation of new malaria risk (\\href{https://www.researchgate.net/publication/320234398_Public_preferences_and_valuation_of_new_malaria_risk}{link}) 
    \\item Attitudes towards public health spending (\\href{https://www.sussex.ac.uk/webteam/gateway/file.php?name=wps-07-2019.pdf&site=24}{link}) 
    \\end{itemize}",
    PhD_info[1], PhD_info[2], PhD_info[3], PhD_info[4],
    "This was financed by an EU FP7 project: HELIX (\\href{https://www.helixclimate.eu/}{link}). This involved multidisciplinary collaboration with researchers from mathematics, physics, engineering, geography and sociology.",
    # MSc
    MSc_info[1], MSc_info[2], MSc_info[3], MSc_info[4], 
      "Specialization: Applied micro-econometrics",
    MSc_info[1], MSc_info[2], MSc_info[3], MSc_info[4], 
      "Thesis: 'What explains willingness to pay for avoiding morbidity risk due to malaria? Results from a global meta-analysis' (later on became first chapter of PhD thesis)",
    # BSc
    BSc_info[1], BSc_info[2], BSc_info[3], BSc_info[4], 
      "Specialization: Academic (i.e. a special focus on applied mathematics and statistics)"
  ) %>% 
  detailed_entries(degree, uni, loc, dates, details)
```
